---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Import the relevant data files

```{r}
library(ECISR)

byserumfree.df = ecis_import_long("HCMVEC/resample.abp", "HCMVEC/model.csv", "HCMVEC/by_treatment.csv")

byallfactors.df = ecis_import_long("HCMVEC/resample.abp", "HCMVEC/model.csv", "HCMVEC/with_growth_discriminated.csv")

bygrowthmedia.df = ecis_import_long("HCMVEC/resample.abp", "HCMVEC/model.csv", "HCMVEC/bygrowthmedia.csv")

```


First we plot out the growth phase as individual replicates

```{r}
bygrowthmediadsubset.df = subset(bygrowthmedia.df, Time<50)
bygrowthmediadsubset.df = subset(bygrowthmediadsubset.df, Sample != "Grafton media only")
bygrowthmediadsubset.df = subset(bygrowthmediadsubset.df, Sample != "SBS media only")
forcm.df = subset(bygrowthmediadsubset.df, Time>1)


ecis_plot_all(bygrowthmediadsubset.df, "R",4000)
ecis_plot_all(bygrowthmediadsubset.df, "Rb",0)
ecis_plot_all(forcm.df, "Cm",0)
ecis_plot_all(bygrowthmediadsubset.df, "Alpha",0)
ecis_plot_all(bygrowthmediadsubset.df, "RMSE",0)

```

Then we plot out a summary

```{r}
ecis_plot_summary(bygrowthmediadsubset.df, "R",4000)
ecis_plot_summary(bygrowthmediadsubset.df, "Rb",0)
ecis_plot_summary(forcm.df, "Cm",0)
ecis_plot_summary(bygrowthmediadsubset.df, "Alpha",0)
ecis_plot_summary(bygrowthmediadsubset.df, "RMSE",0)
```


Now we look at the serum free portion

```{r}

subbyserumfree.df = subset(byserumfree.df, Sample != "Media only")
serumarea.df = subset(subbyserumfree.df, Time>40)

ecis_plot_all(serumarea.df, "R", 4000)
ecis_plot_all(serumarea.df, "Rb", 0)
ecis_plot_all(serumarea.df, "Cm", 0)
ecis_plot_all(serumarea.df, "Alpha", 0)
ecis_plot_all(serumarea.df, "RMSE", 0)



```

Then we normalise it to remove a lot of the variation

```{r}
normalserumarea.df = ecis_normalise(serumarea.df, 50, divide = FALSE)

ecis_plot_all(normalserumarea.df, "R", 4000)
ecis_plot_all(normalserumarea.df, "Rb", 0)
ecis_plot_all(normalserumarea.df, "Cm", 0)
ecis_plot_all(normalserumarea.df, "Alpha", 0)
ecis_plot_all(normalserumarea.df, "RMSE", 0)

````

Then we summarise the thing

```{r}
normalserumarea.df = subset(normalserumarea.df, Time<120)

ecis_plot_summary(normalserumarea.df, "R", 4000)
ecis_plot_summary(normalserumarea.df, "Rb", 0)
ecis_plot_summary(normalserumarea.df, "Cm", 0)
ecis_plot_summary(normalserumarea.df, "Alpha", 0)
ecis_plot_summary(normalserumarea.df, "RMSE", 0)
```



Finaly we look at if the two factors may have interacted by plotting out each well

```{r}
byallfactorssubset.df = subset(byallfactors.df, Time > 40)

ecis_plot_summary(byallfactorssubset.df, "R", 4000)
ecis_plot_summary(byallfactorssubset.df, "Rb", 0)
ecis_plot_summary(byallfactorssubset.df, "Cm", 0)
ecis_plot_summary(byallfactorssubset.df, "Alpha", 0)
ecis_plot_summary(byallfactorssubset.df, "RMSE", 0)

```

Then we normalise the thing to try and reduce the variation

```{r}
normalbyallfactors.df = ecis_normalise(byallfactorssubset.df, 50, divide = FALSE)

ecis_plot_summary(normalbyallfactors.df, "R", 4000)
ecis_plot_summary(normalbyallfactors.df, "Rb", 0)
ecis_plot_summary(normalbyallfactors.df, "Cm", 0)
ecis_plot_summary(normalbyallfactors.df, "Alpha", 0)
ecis_plot_summary(normalbyallfactors.df, "RMSE", 0)
```








