---
title: "vascr - tools for processing vascular impedance data"
author: "James Hucklesby"
date: "03/04/2020"
# output: rmarkdown::html_vignette
# <!-- vignette: > -->
# <!--   %\VignetteIndexEntry{Vascar} -->
# <!--   %\VignetteEngine{knitr::rmarkdown} -->
# <!--   %\VignetteEncoding{UTF-8} -->
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
ECISR is a suite of tools designed to deal with the large quantaties of data generated by Electrical Cell-Substrate Impedance Sensing (ECIS) instruments. It provides tools to import, standardise, plot and output data sets, as well as running statistical analyasis out of the box. The package is designed to be as easy to use as possible, abstracting away much of R's underlying complexity and reducing the barrier to entry for ECIS data analyasis.

# Key design principles
ecisr is built with 3 key design considerations    
* Simplicity - a minimal number of functions should be able to run everything, and arguments should have defaults automatically set where possible
* Visualisation - data should be visualised in an immediatley interpretable way wherever possible    
* Robustness - technical and statistical checks should be run concurrently where possible, and warn the user if they try do do something inappropriate   

# Instalation
```{r eval = FALSE}
# The easiest way to install this package is from CRAN
install.packages("vascr")

# Or the developmetnt version is available from GitHub:
# install.packages("devtools")
devtools::install_github("JamesHucklesby/vascr")

```

```{r}
# In each script, import the library
library(vascr)
```


#File import
Files can be imported using a variety of functions. ECIS and xcelligence data can be imported by directly specifying the native file produced by the instrument, whereas CellZScope data must be exported from the software before being imported. Please see the relevant help functions for specific details.

## Data import functions
ecis_import and cellzscope import can also have _raw or _modeled added to the end of their functions to import just the raw or modeled components of the dataset. This can be used for debugging, or to speed up execution where only one of the datasets are of interest to the experimenter. xcelligence_import is able to access the original data captured by the xcelligence instrument at all three resoltuions, giving better data access than the shipped software and allowing for more direct comparason with other impedance instruments.

```{r eval=FALSE}
ecis_import(rawdata, modeleddata, lookups)
cellzscope_import(rawdata, modeleddata, lookups)
xcelligence_import(rawdata, lookups)
```

## Lookup tables
Lookup tables tell vascr which wells of the 96 well plate contain which samples. This allows 

Lookup tables should be constructed with a column named Well in the top leftmost cell, with well names (in the format A01) in any order beneath them. It is not required that all wells are addressed, as wells that have no lookup specified will be given the sample name "NA". Characteristics such as cell line, number of cells or the type of vechicle should be specified as additional columns. These can be named anything the user wishes, and capitalisation and spaces shoudld be used in the way that they should appear on the final graphic. Below each title, the corresponding number or class should be input. Cells may be left blank, and these will be considered as not containing the specified item. Please see the file used to generate growth.df below. These files can be easily generated in excell and saved as either excell or CSV formats. Where this data may already exist in the data file, such as in a throughly set up xcelligence or cellZScope eperiment, this data can be used by not specifying a lookup file. However, a lookup file is still recomended as it allows for much finer control over the data once imported to vascr.

```{r}
lookup = system.file("/extdata/growth1_samples.csv", package = "vascr")
lookuptable = read.csv(lookup)
head(lookuptable)
```

## Dealing with multiple experiments
Vascr has a built-in understanding of multiple experiments, and the import process reflects this. The import process can be repeated multiple times, as illustrated in the example below. The experimentname argument can be used to manually specify an experiment name, otherwise the file name that the data was derrived from is used. This name shoud be specified as you wish for it to appear in the final graphs and datasets that will be produced.


```{r}
modeled1 = system.file('extdata/growth1_raw_TimeResample_RbA.csv', package = 'vascr')
modeled2 = system.file('extdata/growth2_raw_TimeResample_RbA.csv', package = 'vascr')
modeled3 = system.file('extdata/growth3_raw_TimeResample_RbA.csv', package = 'vascr')

raw1 = system.file('extdata/growth1_raw_TimeResample.abp', package = 'vascr')
raw2 = system.file('extdata/growth2_raw_TimeResample.abp', package = 'vascr')
raw3 = system.file('extdata/growth3_raw_TimeResample.abp', package = 'vascr')

key1 = system.file('extdata/growth1_samples.csv', package = 'vascr')
key2 = system.file('extdata/growth2_samples.csv', package = 'vascr')
key3 = system.file('extdata/growth3_samples.csv', package = 'vascr')

growth1.df = ecis_import(raw1, modeled1, key1, "One")
growth2.df = ecis_import(raw2, modeled2, key2, "Two")
growth3.df = ecis_import(raw3, modeled3, key3, "Three")

# Then we resample these datasets 
growth1_resample.df = vascr_resample(growth1.df, by = 10, from = 0, to=200)
growth2_resample.df = vascr_resample(growth2.df, by = 10, from = 0, to=200)
growth3_resample.df = vascr_resample(growth3.df, by = 10, from=0, to=200)

#Then we combine the datsets
growth = vascr_combine(growth1_resample.df, growth2_resample.df, growth3_resample.df)

```

## Resampling data
Impedance data may not be collected at exactly the same time in evey experiment. This may be due to differing numbers of wells needing to be measured, small changes in the time taken to treat various experiments or minor changes to experimental paramaters. These tiny differences have a neglegible effect on the overall inferences of the data collected, however even slightly different values prevent statistical analyasis from being carried out and makes plotting of the data challenging. Furthermore the time points at which data is collected is often determined by the maximum speed of the electronics within the impedance instrument, rather than any intuative or biologically relevant timing. Althoguh data could be collected at a pre-determined rational frequency and treatments forced to always take exactly the same time, this risks missing short-duration responses that can only be collected by impedance sensing. Therefore the optimal approach is to collect the data at the maximum possible rate, and mathematically transform the data later to make the timecourses absolutley identical. Furthermore this approach allows a frequency to be selected that optimises computational resources by only visualising at an appropriate frequency, and allows for robust direct comparason between experiments and instruments.

Within vascr, resampling by linear interpolation provides the method for this process. Mathematically, the change between data points is assumed to be linear, with a line fitted between each pair of points for each experiment, well and measurment combination. This assumption of linearlity is generaly highly accurate, as impedance instruments generally collect data at a high frequency, and changes occour over relatively long time periods. Furthermore, this interpolation only requires two data points, which means that sudden changes are not averaged out by incorporating a large number of other data points.

The function to carry out this analyasis is vascr_resample. This function requires a vascr dataset to be selected, and the resampling frequency to be specified. For the purposes of this document, we resample every 10 hours as this is sufficent information to generate accurate visualisations without slowing down data analyasis too much. As resamping is a relativley computationally intensive task, maximum and minimum values can also be specified. This allows the user to only resample the section of interest, saving significant computational time and generating a cleaner dataset for further analyasis.


## Combining datasets
Although R contains built in functions for joining data frames end to end, these simply combine the datasets without checking that the dataset as a whole is still valid. vascr_combine confirms that the input datasets conform to the vascr data format, and will provide informative errors if that is not the case. Furthermore it will check that the data is sampled at the same frequency,  and will automatically re-number experiments that have the same name. If the frequencies are not identical, it will warn the user and the function can be repeated with the argument resample = TRUE, which will automatically resample at the lowest frequency specified. 


# Vascr dataset structure

```{r}
head(vascr_subset(growth.df,unit = "R", frequency = 4000))
```

A vascr dataset has the following columns that are compulsary and must be specified

An ECIS data frame has 8 core variables and the ability to have additional user-defined variables added

8 core variables:
Time - Time at which the reading was taken (in hours)  
Well - Which well the reading was taken from  
Unit - Unit value is measured in. See next section for details.
Value - The value of the reading taken  
Sample - The sample name. Multiple wells with the same sample name will be automatically grouped for graphing  
Frequency - The frequency at which the reading was taken. All modeled variables have a frequency of 0  
Experiment - The file name from which the experiment was generated
Instrument - The type of instrument that the data was aquired on

There are also an additional two columns that may be optionally added:
Vehicle - What vehicle the treatment is delivered in. This is used to automatically include vehicles with appropriate wells
Is.Vehicle - TRUE or FALSE. If not specified, this is automatically generated by assigning it to the well for each vehicle with the fewest things added

Other columns will also be added, based on what you stated in your csv when you imported the files. In growth.df these are:
cells - the number of cell initially added per well
line - the cell line added. In the case of the dataset these are all the same, but the data is preserved in case this data is compared to other cell lines

All import functions import data files in this format. This format is required for other vascr funcitons in this package to work correctly.

# Units used in this package
Although impedance analysers all work on the same principle and collect the same raw data, this is then processed in quite different ways. The table below shows which units can be processed, if they are modelled and which instruments collect the particular value. The summary unit used as short hand within the vascr package is also shown. The content of the "content" column is what will be presented in graphs that are automatically labeled, with symbol named being converted into symbols where possible.

```{r}
vascr_units_table()
```


# Data levels
Vascr data can be summarsied in multiple ways, by specifying the level. These work both in vascr_summarsise and vascr_plot functions
structure - returns a list of which well contain which sample
deviation - returns the percent deviaion of each well at each time point
max_deviation - returns the maximum deviation of each well at any time point
wells - returns individual data for each well at each data point
experiments - returns the mean data for each sample in each experiment
summary - returns the mean value of the experimental mean values for the dataset overall. The statistics are set up so as to display the inter-experimental, not intra-experimental variation, giving a more accurate display of experimental variation accross the population

```{r}
vascr_plot(growth1_resample.df, level = "structure")
vascr_plot(growth1_resample.df, level = "deviation")
vascr_plot(growth1_resample.df, level = "max_deviation")
vascr_plot(growth, level = "wells")
vascr_plot(growth, level = "experiments")
vascr_plot(growth, level = "summary")

```

## Data cleaning

```{r}
vascr_plot(growth.df, unit = "R", frequency = 4000, level = "deviation")
```



```{r}
# Plot well quality heatmap (or actually, could use the multiplot infastructure to do this many times (one per expt). That's a good thought. Same for plot_plate actually)

# We remove growth1 well D1 (and another test well to demonstrate we can) because it is defective. We also remove the 0 cells as it is not sensible to include
vascr_plot_isolate(growth1_resample.df, well = "D01")

growth1_resample.df=  vascr_exclude(growth1_resample.df, wells = c("D01", "I04"), sample = "0_cells + HCMEC D3_line")

vascr_plot(growth1_resample.df, replication = "plate")
```


# Generating graphics
vascr has a wide variety of plots built in. 

## Selecting frequencies and units
vascr functions have built in arguments for selecting the unit and frequency to plot. This avoids the need to pre-process the data before plotting it, reducing the number of 

```{r}
wells = vascr_plot(growth.df, unit = "Rb", level = "wells", title = "Individual wells")
experiments = vascr_plot(growth.df, unit = "Rb", level = "experiments", title = "Individual experiments")
summary = vascr_plot(growth.df, unit = "Rb", level = "summary", title = "All experiments combined")
summary2 = vascr_plot(growth.df, unit = "Rb", level = "summary", title = "All experiments combined, with traditional error bars", error = 1)

print(vascr_make_panel(wells, experiments, summary, summary2))
```

## Normalising data
Normalisation provides a usefull tool for analysing impedance data, as it removes the variation that occours from differing growth patterns of the cells, allowing the effects of any treatments applied to be more easily visualised. Two methods are available in vascr, subtraction (the default) and division. Both these methods work by selecting the values at the time point that values are normalised to, and then dividing or subtracting this value from all other time points. Subtraction still shows the magnitude of the shift, while division shows fold shifts. Division is able to resolve smaller differences, however it is more strongly affected by outliers so may make differences appear large, even though they are not statistically or practically significant.

vascr plot can also call vascr_normalise from within it's arguements, saving the user from having to generate this data separatley. This is done using the normtime and divide arguments directly within the vascr_plot call.

```{r}
#Show that a single key time point can be plotted
vascr_plot(growth, time = 50, replication = "summary")
vascr_plot(growth, time = 50, replication = "experiments")
vascr_plot(growth, time = 50, replication = "wells")
```



```{r}
# Data can be normalised by division or subtraction
subtract = vascr_plot(growth.df, normtime = 100, title = "Normalisation by subtraction") # Subtraction is the default
divide = vascr_plot(growth.df, normtime = 100, divide = TRUE, title =  "Normalisation by division")

print(vascr_make_panel(subtract, divide))
```


# Subsetting continuous variables


# Plotting continuous variables
```{r}
wellscont = vascr_plot(growth.df, time = 100, visualisation = "bar")



wellscont = vascr_plot_continuous(growth.df, unit = "R", frequency = 4000, level = "wells", time = 100, error = Inf, priority = c("cells", "Experiment"))
experimentscont = vascr_plot_continuous(growth.df, unit = "R", frequency = 4000, level = "experiments", time = 100, priority = c("cells", "Experiment"), error = 1)
summarycont = vascr_plot_continuous(growth.df, unit = "R", frequency = 4000, level = "summary", time = 100, continuous = "cells", error = Inf, priority = c("cells", "Experiment"))



print(vascr_make_panel(wells, experiments, summary))
```

```{r}
p1 = vascr_plot_bar(growth.df, level = "wells", frequency = 4000, unit = "R", time = 50, error = Inf)
p2 = vascr_plot_bar(growth.df, level = "experiments", frequency = 4000, unit = "R", time = 50)
p3 = vascr_plot_bar(growth.df, level = "summary", frequency = 4000, unit = "R", time = 50, error = Inf)

p4 = vascr_plot_continuous(growth.df, unit = "R", frequency = 4000, level = "wells", time = 100, error = Inf, priority = c("cells", "Experiment"))
p5 = vascr_plot_continuous(growth.df, unit = "R", frequency = 4000, level = "experiments", time = 100, priority = c("cells", "Experiment"), error = 1)
p6 = vascr_plot_continuous(growth.df, unit = "R", frequency = 4000, level = "summary", time = 100, continuous = "cells", error = Inf, priority = c("cells", "Experiment"))

vascr_make_panel(p1, p2, p3, p4, p5, p6)
```



# Representaiton of error bars

Add dots to the code

```{r}
# Demonstrate that error can be changed
single = vascr_plot(growth.df, error = 1)
subsampled = vascr_plot(growth.df, error = 5)
no = vascr_plot(growth.df, error = 0)
continuous = vascr_plot(growth.df, error = Inf)

```

# Styling graphs
The graphs generated by vascr are ggplots, and can be styled or added to using standard ggplot statements. Some of these styling options are also built into vascr_plot for ease of use. The line thickness can be adjusted wtih the linesize attrubute. The alpha value of ribbon fill can be  ajusted using the alphavalue variable. Higher alpha values make errors more visble, whilst lower values can help to decultter busy graphs.

```{r}
vascr_plot(growth.df) + ylim(0,750)

vascr_plot(growth, linesize = 5)
vascr_plot(growth, alphavalue = 0.8)

```


```{r}
# Tiles can be customised, although they are automatically generated
vascr_plot(growth, title = "Growth of HCMEC/D3 cells", xlab = "Time in hours", ylab = "Resistance in ohm")
```


```{r}
# Demonstrate that only some samples can be plotted, or that the default name-trimming behaviour can be turned off
vascr_plot(growth, samplecontains = "5,000")
vascr_plot(growth, stripidentical = FALSE)
```


```{r}
# Show that we can zoom in on a time area if needed, and that this time can be a variable
vascr_plot(growth, time = 100)
vascr_plot(growth, time = c(50,100))
```



```{r}

# Demonstrate that arguements are cumulative
vascr_plot(growth.df, level = "wells", linesize = 3, error = 5, normtime = 100)

```


```{r}
# Show some of the pre-prepared matrices for lots of data
vascr_multiplot(data = growth.df, frequency = 4000, time = list(c(50, 100),100), replication = "wells", unit = "Rb")
vascr_multiplot(data = growth.df, frequency = 4000, time = list(50, 100))


vascr_plot(growth.df, "R")
vascr_plot_model(growth)
vascr_plot_model(growth, error = 1) # You can also pass vascr_plot arguements through if you want to affect how the plot looks
```

#Add statistics to the plot

```{r}
vascr_plot(growth, time = 50, replication = "summary", confidence = .05)

# Run ANOVA analyasis
vascr_plot_anova(growth.df, time = 50, unit = "R", frequency = 4000)
vascr_plot_anova(growth.df, time = 100, unit = "R", frequency = 4000)
```






















