---
title: "vascr - tools for processing vascular impedance data"
author: "James Hucklesby"
date: "03/04/2020"
# output: rmarkdown::html_vignette
# <!-- vignette: > -->
# <!--   %\VignetteIndexEntry{Vascar} -->
# <!--   %\VignetteEngine{knitr::rmarkdown} -->
# <!--   %\VignetteEncoding{UTF-8} -->
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
ECISR is a suite of tools designed to deal with the large quantaties of data generated by Electrical Cell-Substrate Impedance Sensing (ECIS) instruments. It provides tools to import, standardise, plot and output data sets, as well as running statistical analyasis out of the box. The package is designed to be as easy to use as possible, abstracting away much of R's underlying complexity and reducing the barrier to entry for ECIS data analyasis.

# Key design principles
ecisr is built with 3 key design considerations    
* Simplicity - a minimal number of functions and arguements should be able to run everything    
* Visualisation - data should be visualised wherever possible    
* Robustness - technical and statistical checks should be run continuously, and warn the user if they try do do something inappropriate   


## Instalation
```{r eval = FALSE}
# The easiest way to install this package is from CRAN
install.packages("vascr")

# Or the developmetnt version is available from GitHub:
# install.packages("devtools")
devtools::install_github("JamesHucklesby/vascr")

```

```{r}
# In each script, import the library
library(vascr)
```


## File import
Files can be imported using a variety of functions. ECIS and xcelligence data can be imported by directly specifying the native file produced by the instrument, whereas CellZScope data must be exported from the software before being imported. Please see the relevant help functions for specific details.

ecis_import and cellzscope import can also have _raw or _modeled added to the end of their functions to import just the raw or modeled components of the dataset. This can be used for debugging, or to speed up execution where only one of the datasets are of interest to the experimenter. xcelligence_import is able to access the original data captured by the xcelligence instrument at all three resoltuions, giving better data access than the shipped software and allowing for more direct comparason with other impedance instruments.

```{r eval=FALSE}
ecis_import(rawdata, modeleddata, lookups)
cellzscope_import(rawdata, modeleddata, lookups)
xcelligence_import(rawdata, lookups)
```

Lookup tables should be constructed with a column named Well in the top leftmost cell, with well names (in the format A01) in any order beneath them. It is not required that all wells are addressed, as wells that have no lookup specified will be given the sample name "NA". Characteristics such as cell line, number of cells or the type of vechicle should be specified as additional columns. These can be named anything the user wishes, and capitalisation and spaces shoudld be used in the way that they should appear on the final graphic. Below each title, the corresponding number or class should be input. Cells may be left blank, and these will be considered as not containing the specified item. Please see the file used to generate growth.df below. These files can be easily generated in excell and saved as either excell or CSV formats.

```{r}
lookup = system.file("/extdata/growth1_samples.csv", package = "vascr")
lookuptable = read.csv(lookup)
head(lookuptable)
```

## Dealing with multiple experiments
Vascr has a built-in understanding of multiple experiments, and the import process reflects this. The import process can be repeated multiple times, as illustrated in the example below. The experimentname argument can be used to manually specify an experiment name, otherwise the file name that the data was derrived from is used. Again, this shoud be specified as you wish for it to appear in the final dataset.

# Resampling data
Vascr data can be resampled, meaning that the datasets are linearly interpolated and new datapoints are generated from that set. This is important, as it allows for data points to be collected at the same time and therefore correctly compared. Although this is built into some instruments, some will have slight differences due to factors such as how many wells are filled or how treatment was preformed. Resampling remves this variation. Furthermore this allows for plots to be generated with accurate error bars at round intervals rather than uneven ones, making the data easier to interpret.

```{r}
modeled1 = system.file('extdata/growth1_raw_TimeResample_RbA.csv', package = 'vascr')
modeled2 = system.file('extdata/growth2_raw_TimeResample_RbA.csv', package = 'vascr')
modeled3 = system.file('extdata/growth3_raw_TimeResample_RbA.csv', package = 'vascr')

raw1 = system.file('extdata/growth1_raw_TimeResample.abp', package = 'vascr')
raw2 = system.file('extdata/growth2_raw_TimeResample.abp', package = 'vascr')
raw3 = system.file('extdata/growth3_raw_TimeResample.abp', package = 'vascr')

key1 = system.file('extdata/growth1_samples.csv', package = 'vascr')
key2 = system.file('extdata/growth2_samples.csv', package = 'vascr')
key3 = system.file('extdata/growth3_samples.csv', package = 'vascr')

growth1.df = ecis_import(raw1, modeled1, key1, "One")
growth2.df = ecis_import(raw2, modeled2, key2, "Two")
growth3.df = ecis_import(raw3, modeled3, key3, "Three")

# Then we resample these datasets 
growth1_resample.df = vascr_resample(growth1.df, 10, 0, 180)
growth2_resample.df = vascr_resample(growth2.df, 10, 0, 200)
growth3_resample.df = vascr_resample(growth3.df, 10, 0, 200)

# Then we check that the samples look ok
vascr_plot(growth1_resample.df, replication = "plate")
vascr_plot(growth2_resample.df, replication = "plate")
vascr_plot(growth3_resample.df, replication = "plate")
```

## Vascr dataset structure

```{r}
head(growth.df)
```

A vascr dataset has the following columns that are compulsary and must be specified

An ECIS data frame has 8 core variables and the ability to have additional user-defined variables added

8 core variables:
Time - Time at which the reading was taken (in hours)  
Well - Which well the reading was taken from  
Unit - Unit value is measured in. See next section for details.
Value - The value of the reading taken  
Sample - The sample name. Multiple wells with the same sample name will be automatically grouped for graphing  
Frequency - The frequency at which the reading was taken. All modeled variables have a frequency of 0  
Experiment - The file name from which the experiment was generated
Instrument - The type of instrument that the data was aquired on

There are also an additional two columns that may be added:
Vehicle - What vehicle the treatment is delivered in. This is used to automatically include vehicles with appropriate wells
Is.Vehicle - TRUE or FALSE. If not specified, this is automatically generated by assigning it to the well for each vehicle with the fewest things added

Other columns will also be added, based on what you stated in your csv when you imported the files. In growth.df these are:
cells - the number of cell initially added per well
line - the cell line added. In the case of the dataset these are all the same, but the data is preserved in case this data is compared to other cell lines

All import functions import data files in this format. This format is required for other ecis funcitons in this package to work correctly.

# Units used in this package
```{r}
vascr_units_table()
```




```{r}
vascr_plot_plate(growth1_resample.df, replication = "plate")
vascr_plot(growth2_resample.df, replication = "plate")
vascr_plot(growth3_resample.df, replication = "plate")
```

## Data cleaning

```{r}
vascr_plot_quality_plate(growth1.df)
```



```{r}
# Plot well quality heatmap (or actually, could use the multiplot infastructure to do this many times (one per expt). That's a good thought. Same for plot_plate actually)

# We remove growth1 well D1 (and another test well to demonstrate we can) because it is defective. We also remove the 0 cells as it is not sensible to include
vascr_plot_isolate(growth1_resample.df, well = "D01")

growth1_resample.df=  vascr_exclude(growth1_resample.df, wells = c("D01", "I04"), sample = "0_cells + HCMEC D3_line")

vascr_plot(growth1_resample.df, replication = "plate")
```


# Combining datasets
Although R contains built in functions for joining data frames end to end, these simply combine the datasets without checking that the dataset as a whole has integrity. vascr_combine confirms that the input datasets conform to the vascr data format, and will provide intelegent errors if that is not the case. Furthermore it will check that the data is sampled at the same frequency,  and will automatically re-number experiments that have the same name. If the frequencies are not identical, it will warn the user and the function can be repeated with the argument resample = TRUE, which will automatically resample at the lowest frequency specified. 

```{r}
#Then we combine the datsets
growth = vascr_combine(growth1_resample.df, growth2_resample.df, growth3_resample.df)
```


# Replication levels used in this package

Replication
all = plots each replicate from all experiments in a spaghetti graph
wells = plots out each well in it's own seperate facet of a large graph
experiment = plots each experiment as a single experiment
summary = plots the summary data from multiple independent experiments. The statistics are set up so as to display the inter-experimental, not intra-experimental variation, giving a more accurate display of experimental variation accross the population

```{r}
# Demonstrate the three resolutions we can use to display data
vascr_plot(growth, replication = "wells")
vascr_plot(growth, replication = "experiments")
vascr_plot(growth, replication = "summary")
```

```{r}
#Show how to select different units and frequencies
vascr_plot(growth, unit = "Rb")
vascr_plot(growth, unit = "R", frequency = 4000)
```

```{r}
# Data can be normalised by division or subtraction
vascr_plot(growth, normtime = 100) # Subtraction is the default
vascr_plot(growth, normtime = 100, divide = TRUE)
```

```{r}
#Show that a single key time point can be plotted
vascr_plot(growth, time = 50, replication = "summary")
vascr_plot(growth, time = 50, replication = "experiments")
vascr_plot(growth, time = 50, replication = "wells")
```

# Subsetting continuous variables


# Plotting continuous variables


# Representaiton of error bars

Add dots to the code

```{r}
# Demonstrate that error can be changed
vascr_plot(growth, error = 1)
vascr_plot(growth, error = 5)
vascr_plot(growth, error = 0)
vascr_plot(growth, error = Inf)

```


```{r}
# Demonstrate some style changes
vascr_plot(growth, linesize = 5)
vascr_plot(growth, alphavalue = 0.8)

vascr_multiplot(data = growth.df, frequency = 4000, time = list(c(50, 100),100), replication = "wells", unit = "Rb")

 vascr_multiplot(data = growth.df, frequency = 4000, time = list(50, 100))

vascr_plot(growth.df) + ylim(0,750)

```


```{r}
# Tiles can be customised, although they are automatically generated
vascr_plot(growth, title = "Growth of HCMEC/D3 cells", xlab = "Time in hours", ylab = "Resistance in ohm")
```


```{r}
# Demonstrate that only some samples can be plotted, or that the default name-trimming behaviour can be turned off
vascr_plot(growth, samplecontains = "5,000")
vascr_plot(growth, stripidentical = FALSE)
```


```{r}
# Show that we can zoom in on a time area if needed, and that this time can be a variable
vascr_plot(growth, time = 100)
vascr_plot(growth, time = c(50,100))
```



```{r}

# Demonstrate that arguements are cumulative
vascr_plot(growth, replication = "wells", linesize = 3, error = 5, normtime = 100)

```


```{r}
# Show some of the pre-prepared matrices for lots of data
vascr_plot_spectra(growth, "R")
vascr_plot_model(growth)
vascr_plot_model(growth, error = 1) # You can also pass vascr_plot arguements through if you want to affect how the plot looks
```


```{r}
#Add statistics to the plot
vascr_plot(growth, time = 50, replication = "summary", confidence = .05)

# Run ANOVA analyasis
vascr_ANOVA(growth, time = 50, unit = "R", frequency = 4000)
vascr_ANOVA(growth, time = 100, unit = "R", frequency = 4000)
```






















