---
title: "vascr - a step by step guide"
author: "James Hucklesby"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vascr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

ECISR is a suite of tools designed to deal with the large quantaties of data generated by Electrical Cell-Substrate Impedance Sensing (ECIS) instruments. It provides tools to import, standardise, plot and output data sets, as well as running statistical analysis out of the box. The package is designed to be as easy to use as possible, abstracting away much of R's underlying complexity and reducing the barrier to entry for ECIS data analysis.

# Key design principles
ecisr is built with 3 key design considerations    
* Simplicity - a minimal number of functions and arguements should be able to run everything    
* Visualisation - data should be visualised wherever possible    
* Robustness - technical and statistical checks should be run continuously, and warn the user if they try do do something inappropriate   

## A note on why ecisr is superior to using the built in ABP software
Although the Applied Biophysics software produces graphs and is adequate for basic data analysis, it lacks the ability to make publication ready graphs and run statistical analyasis. It also stores the data internaly in a proprietary format, making it dificult to extend and directly work with the dataset.    

By contrast ECISR allows direct access to the data using R, and has statistical analysis tools built in. The graphs ECISR generates are standard ggplot2 objects, meaning that you can modifiy them further in R. Alternatively, ECISR will allow you to output your data in a prism friendly format, allowing for graphing elsewhere without having to first run the data through excel. ECISR also contains tools to run ANOVA analyasis, and is able to select an appropriate statistical analysis automatically by inspecting the dataset. R's bulit in statistical tools can also be employed, allowing for custom statistical analyasis as needed.

## Analyasis structure
ECIS data visualisation is best conducted in 4 stages:   
* Import of raw data   
* Cleaning and removal of erronious data   
* Plotting and interpretation of results   
* Formal statistical testing   

We will follow these steps in this document, using a test dataset, growth.df. This dataset contains three independent ECIS experiments plotting the growth of HCMEC/D3 cells at various seeding densities.

# Import the dataset to R


### Prepare the data using Applied Biophysics ECIS software
Firstly, we need to use the modeling tools built into the ECIS software to model the data. Skip this step if you want to work with 

To prepare data to be imported you must do the following:     
1. Open the datafile to be analised in the ABP ECIS software       
2. Ensure that the offset slider is set to the minimum position, the range slider is set to the maximum position and all wells are selected so all the data is visible. I do not recomend that any data is excluded at this point (including cell free and replicates that are obveously going to be excluded) as removing the wells later in R is more traceable, and you can comment in why you have chosen to remove the well for reference.      
3. Select the "Model" radio button in the right hand pannel of the ECIS software   
4. Press the  "Model" button   
5. Select the correct cell free reference radio button and press model   
6. Once completed (this may take some time), press "Save RbA" button to save down the modelled data CSV document.   


Then generate a CSV file in the following format. These are the final names that will end up on the graphs generated from this dataset. 

Well  | cell number | cell line
------------- | -------------
A1  | 5000 |HCMEC D3
A2  | 10,000 |HCMEC D3

Repeat this process for each of the replicate experiments. You should end up with three files for each experiment, saved in a folder on your computer.

### Create a new R project in RStudio, and install ECISR

For the purposes of this article, we will use RStudio for data analyasis. This is an IDE that allows us to work with R in a nicer enviroment, and keep track of various internal variables. It is advisable to set up a new directory for each new project, 

1. Install and open RStudio (rstudio.com or via your institutions software portal). This sofware is available free of charge.
2. Select File, New Project
3. Select Existing Directory
4. Select the directory containing the files you created in the previous step, then press "Create Project"

### Import the files to ECISR

We must first install and load in the ECISR package. This loads up all the tools we need to analyse our dataset.
```{r eval = FALSE}
# First we install devtools, which allows us to then install the package from GitHub. Then we use it to install ecisr
#This code should be run once, when you start using the package
if (!require(devtools)) install.packages('devtools')
require(devtools)
if (!require(ecisr))     devtools::install_github("JamesHucklesby/ECIS-R", auth_token == "b8b71100e6dfcb467c3777abacafbccd0dbfec94", ref = "master")
```


Firstly, we import the files individually before merging them into a single file. The combining function also checks that the data is in the correct format, and that the files are equivelent.

```{r eval = FALSE}
library(ecisr)

# First we import all 3 datasets to R (commented out here to make package generation faster)
growth1.df = ecis_import("growth1_raw.abp", "growth1_modeled.csv", "growth1_samples.csv")
growth1resample = ecis_resample(growth1.df, 5, 0,200)

growth2.df = ecis_import("growth2_raw.abp", "growth2_modeled.csv", "growth2_samples.csv")
growth2resample = ecis_resample(growth2.df, 5, 0,200)

growth3.df = ecis_import("growth3_raw.abp", "growth3_modeled.csv", "growth3_samples.csv")
growth3resample = ecis_resample(growth3.df, 5, 0,200)


ecis_plot(resampled)

head(growth.df)
```
```{r}

```


The two smaller functions ecis_import_raw (rawdata.abp, lookups) or ecis_import_modelled(modelled.csv, lookups) can be used if only the raw or moddeled data respectivley are required.

## Structure of an ECIS data frame

An ECIS data frame has 8 core variables and the ability to have additional user-defined variables added

Time - Time at which the reading was taken (in hours)  
Well - Which well the reading was taken from  
Unit - Unit value is measured in. Options are Rb, Alpha, Cm, RMSE, Drift, R, X, Z, C, P  
Value - The value of the reading taken  
Sample - The sample name. Multiple wells with the same sample name will be automatically grouped for graphing  
Frequency - The frequency at which the reading was taken. All modeled variables have a frequency of 0  
Experiment - The file name from which the experiment was generated  

other columns will also be added, based on what you stated in your csv when you imported the files

ecis_import imports data files in this format. This format is required for other ecis funcitons in this package to work correctly.

## Resample the data
ECIS generates very large ammounts of data. Furthermore, different experiments are often sampled by the ECIS machine at slightly different intervals, making direct comparsons dificult. To fix this problem, we resample the data so that the time points within it are sensible.

```{r eval = FALSE}
growth2.df = ecis_resample(ecisr::growth.df, 1, from = 0, to = 200)
ecis_plot(growth2.df)

```


## ecis_plot (data.df, )

This function is the main plotting worker function for R. It is a recapitulation of the ggplot2 package, allowing for the easy and rapid generation of plots. This in turn abstracts away much of the complexity of the pre-analyasis of using ggplot, allowing for easier generation of graphs.


levels
all = plots each replicate from all experiments in a spaghetti graph
wells = plots out each well in it's own seperate facet of a large graph
experiment = plots each experiment as a single experiment
summary = plots the summary data from multiple independent experiments. The statistics are set up so as to display the inter-experimental, not intra-experimental variation, giving a more accurate display of experimental variation accross the population

time
Time is set automaticaly to infinity. This triggers the generation of a line plot showing the change in a variable over time. By nominating a time (in hours) a bar plot of the differneces at that time point will be generated. This "timeslice" graph is usefull in evaluating the output of ANOVA and subsequent statistical analyasis.

unit
