---
title: "vascr - tools for processing vascular impedance data"
author: "James Hucklesby"
date: "03/04/2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vascar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
vascr provides a toolbox for the analyasis of data generated by ECIS, cellZscope and xcelligence systems. Functions are provided for importing data from each instrument, as well as rapidly running common processing tasks and generating graphical and statistical representations of the data.

## Instalation
```{r eval = FALSE}
# The easiest way to install this package is from CRAN
install.packages("vascr")

# Or the developmetnt version is available from GitHub:
# install.packages("devtools")
devtools::install_github("JamesHucklesby/vascr")


```


## File import
Files can be imported using a variety of functions. ECIS and xcelligence data can be imported by directly specifying the native file produced by the instrument, whereas CellZScope data must be exported from the software before being imported. Please see the relevant help functions for specific details.

ecis_import and cellzscope import can also have _raw or _modeled added to the end of their functions to import just the raw or modeled components of the dataset. This can be used for debugging, or to speed up execution where only one of the datasets are of interest to the experimenter. xcelligence_import is able to access the original data captured by the xcelligence instrument at all three resoltuions, giving better data access than the shipped software and allowing for more direct comparason with 

```{r}
ecis_import()
cellzscope_import()
xcelligence_import()
```




