---
title: "ECISR"
author: "James Hucklesby"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

ECISR is a suite of tools designed to deal with the large quantaties of data generated by Electrical Cell-Substrate Impedance Sensing instruments. It provides tools to import, standardise, plot and output data sets, as well as running statistical analyasis out of the box. The package is designed to be as easy to use as possible, abstracting away much of R, reducing the barrier to entry for ECIS data analyasis.


## ecis_import (resampled.abp, modelled.csv, lookups.csv)

To prepare data to be imported you must do the following:
Open the datafile to be analised in the APP ECIS software
Select "Help > Show Expert Toolbar Items"
Select "Edit > Resample Time". All datasets to be combined must be resampled at the same frequency. (this may be automated in a later version of this software).
Select the "Model" button in the right hand pannel of the ECIS software. Model all wells as normal.
Once completed, press "Save RbA" button to save down the RbA file

Then generate a CSV file containing the wells used and a sample name. These are the final names that will end up on the graphs generated from this dataset. All files must be in the working directory of the R project. An example of this file is available as part of this software package.

ecis_import will then take all 3 of these files and combine them into a single R dataframe.

The two smaller functions ecis_import_raw (rawdata.abp, lookups) or ecis_import_modelled(modelled.csv, lookups) can be used if only the raw or moddeled data respectivley are required.

## Structure of an ECIS data frame

An ECIS data frame has 8 variables

Time - Time at which the reading was taken (in hours)
Well - Which well the reading was taken from
Value - The value of the reading taken
Sample - The sample name. Multiple wells with the same sample name will be automatically grouped for graphing
Frequency - The frequency at which the reading was taken. All modeled variables have a frequency of 0
Experiment - The file name from which the experiment was generated

ecis_import imports data files in this format. This format is required for other ecis funcitons in this package to work correctly.

## ecis_combine (data1, data2 ...)

ECIS combine returns a data frame contianing both the datasets fed to it. 

It will also check that the time points are resampled to the same frequency, if not it will give a warning, in which case the data frames should be resampled and re-modelled again.

## ecis_plot (data.df, )

This function is the main plotting worker function for R. It is a recapitulation of the ggplot2 package, allowing for the easy and rapid generation of plots. This in turn abstracts away much of the complexity of the pre-analyasis of using ggplot, allowing for easier generation of graphs.


levels
all = plots each replicate from all experiments in a spaghetti graph
experiment = plots each experiment as a single experiment
summary = plots the summary data from multiple independent experiments. The statistics are set up so as to display the inter-experimental, not intra-experimental variation, giving a more accurate display of experimental variation accross the population.

time
Time is set automaticaly to infinity. This triggers the generation of a line plot showing the change in a variable over time. By nominating a time (in hours) a bar plot of the differneces at that time point will be generated. This "timeslice" graph is usefull in evaluating the output of ANOVA and subsequent statistical analyasis.

unit





Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
